\large{\bf Large effects and the infinitesimal model }

### 2023-12-04 writeup.tex
Working on Section 4.

The Gaussian distribution appears thanks to ``the'' Central Limit Theorem,
i.e., the universality of adding up a great many very small and independent things.
However, there is a wider class of Central Limit Theorems,
which describe the distributions of sums of a great many independent and exchangeable things:
the \emph{stable laws} \citep{gnedenko1968limit}.
Roughly speaking, these say that if we add together $n$ independent copies of a random variable $X$
for which $\P\{ X > t \}$ is proportional to $t^{-\alpha}$ for some $0 < \alpha < 2$,
then -- almost regardless of the details -- the result, scaled by $n^{1/\alpha}$,
is well-approximated by a universal form, an $\alpha$-stable distribution
(and the approximation becomes exact, as $n \to \infty$).
Oth the other hand, if $\alpha \ge 2$ then the distribution has finite variance, \revpoint{2}{1} 
and so the sum of $n$, divided by $\sqrt{n}$, converges to a Gaussian distribution.
In such ensembles, single large entries can be important:
concretely, the largest of the $n$ copies will be of order $n^{1/\alpha}$,
so that e.g., for $\alpha=1$ a positive proportion of the sum of $n$ entries will be contributed by a single one.
On the other hand, the largest of $n$ Gaussians is of order $\sqrt{\log n}$. \revpoint{1}{11}

@@

Since the Mendelian sampling term is a sum of terms of the form
$$
    (X_\ell - 1/2) \eta_\ell^{[1]} + (1/2 - X_\ell) \eta_\ell^{[2]},
$$
where $\eta_\ell^{[i]}$ is the effect of the allele in parent $i$ at locus $\ell$,
the appropriate central limit theorem for the sum is determined by
the tail behavior of $|(X - 1/2) \eta|$,
where $X$ is Bernoulli(1/2) and $\eta$ is the effect
of a uniformly chosen one of the two alleles from a uniformly chosen parent
at a uniformly chosen locus.
Since $|X-1/2| = 1/2$, we'd like to know
if $\P\{|X|>2t\}$ is proportional to $t^{-\alpha}$,
i.e., the value of
$$
    \lim_{t \to \infty} \frac{ - \log \P\{ |X| > 2 t \} }{ \log t } = \alpha ,
$$
if the limit exists.
Let $p_\ell$ be the minor allele frequency at the $\ell^\text{th}$ locus;
since the data we have assigns effect 0 to the major allele,
$\eta_\ell = 0$ with probability $1-p_\ell$.
Therefore, if we have estimated effects for $M$ SNPs
with minor allele frequency $p_\ell$ and effect size $e_\ell$ for the minor allele,
for $1 \le \ell \le M$,
then
$$
    \P\{ |\eta| > 2t \} \approx \frac{1}{M} \sum_{\ell : |e_\ell| > 2t} p_\ell .
$$
Note that the relevant effect size distribution here is weighted by frequency, \revpoint{1}{3}
which is necessary because we're looking at the deviation of a typical offspring from their parents,
and the chance a given allele is carried by a random parent is given by the frequency.
(This weighting will make a difference, as frequencies and effect sizes are often correlated in practice.)
So, to obtain a per-trait estimate of $\alpha$,
we plot $\log \frac{1}{M} \sum_{\ell : |e_\ell| > t} p_\ell$ against $\log t$
for a sequence of values of $t$,
and fit a linear model to the tail.
(To do this,
we take as the ``tail''
all values of $t$ for which the number of SNPs with absolute effect above $t$
is at least 20 and at most 10\% of all SNPs.) \revpoint{2}{3}
Examples are shown in Figure~\ref{fig:example_snps}.

@@

Concretely, consider a Moran process where an individual whose trait value is $Z$ dies at rate $\mu(Z)$
and is replaced by the offspring of two randomly chosen individuals,
i.e., a new individual with trait \llabel{offspring_eqn}
\begin{align*}
  \frac{Z^{1} +Z^{2}}{2} + R ,
\end{align*}
where $Z^{1}$ and $Z^{2}$ are independent draws from the empirical population distribution
and $R$ is an independent draw from the Mendelian sampling distribution
(independent of $Z^{1}$ and $Z^{2}$, with distribution to be specified). 
(Note that mutation might enter into $R$, but as the underlying genetic basis is not specified,
we do not attempt to disambiguate the various possible contributions.) \llabel{mutation_note}
Mutation-selection balance demands that that
the distribution of individuals removed by death
and the distribution of individuals added by birth are equal,
\ie that $(Z^{1} +Z^{2})/2 + R$ is equal in distribution
to a $\mu$-weighted choice of $Z$.
% In other words, for all bounded and continuous functions $f$,
% \begin{equation} \label{eqn:MSB_condition}
%     \E\left[ f\left(\frac{Z^{1} +Z^{2}}{2} + R \right) \right] = \frac{\E\left[ f(Z) \mu(Z) \right]}{\E[\mu(Z)]},
% \end{equation}
% where the expectation is with respect to the joint law of the population empirical distribution (the law of $Z$, $Z^{1}$ and $Z^{2}$), and $R$.
% 
% Since the set of functions $\mathcal{S} = \{e^{i \theta x} : \theta \in \R\}$ is separating (see \eg \S 3.4 in \cite{Ethier+Kurtz1986}), it suffices to show that \eqref{eqn:MSB_condition} holds for $f \in \mathcal{S}$, \ie
If the population is large enough, \llabel{deterministic_point}
we can treat the trait distribution as changing deterministically.
Taking the characteristic function of both these distributions,
deterministic mutation-selection balance is therefore equivalent to
\begin{align} \label{eqn:MSB_CF}
    \frac{\E\left[ e^{i \theta Z}\mu(Z) \right]}{\E[\mu(Z)]}
    =
    \E\left[ e^{i\theta\left(\frac{Z^{1} +Z^{2}}{2} + R \right)}\right] .
\end{align}
Now, we assumed that $Z^{1}$, $Z^{2}$, and $R$ are independent, so we can write \eqref{eqn:MSB_CF} as 
\begin{align*}
    \frac{\E\left[ e^{i\theta Z} \mu(Z) \right]}{\E[\mu(Z)]} &= \E\left[ e^{i\theta\left(\frac{Z^{1} +Z^{2}}{2} + R \right)}\right]\\
	% &= \E\left[ e^{i\frac{\theta}{2} Z^{1}}e^{i\frac{\theta}{2} Z^{2}}e^{\i\theta R}\right]\\
	&= \E\left[e^{i\frac{\theta}{2} Z^{1}}\right]\E\left[e^{i\frac{\theta}{2} Z^{2}}\right]\E\left[e^{i\theta R}\right]\\
	&= \phi\left(\frac{\theta}{2}\right)^{2} \Phi(\theta),
\end{align*}
where
$ \phi(\theta) = \mathbb{E}\left[e^{i\theta Z^{1}}\right] = \mathbb{E}\left[e^{i\theta Z^{2}}\right] $
and
$ \Phi(\theta) = \mathbb{E}\left[e^{i\theta R}\right] $
are the characteristic functions of $Z^{j}$ and $R$, respectively. 
In what follows, we will explore some non-Gaussian solutions $\phi(\theta)$ to equation \eqref{eqn:MSB_CF}:

@@

First, consider the case without selection, so $\mu$ is a constant
which we may take equal to 1. 
(To remind the reader, our goal is therefore to describe \llabel{clarification1}
what forms the population distribution might take in a neutral model
if we assume only that the Mendelian sampling terms are independent of parental values,
and nothing else about the distribution of these terms or the initial population distribution.) 
Then, our condition on the characteristic function simplifies to
\begin{equation}\label{eq:CF_neutral}
	{\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta) %e^{-\frac{\eta^{2}\theta^{2}}{2}} 
	= \phi(\theta) ,
\end{equation}
\ie $\phi(\theta)$ is a fixed point of $F$, where
 \[
	F(\phi)(\theta) = {\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta),
\]
whereas any fixed point must be the limit of an arbitrary characteristic function, $\varphi(\theta)$, under repeated applications of $F$,
\ie \revpoint{2}{7}
\begin{align*}
	\phi(\theta) 
    &= \lim_{n \to \infty} F^{(n)}(\varphi)(\theta) \\
    &= \lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=0}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}} .
\end{align*}
% Iterating $F$ $n$ times, we get
% \[
% 	F^{(n)}(\varphi)(\theta) =  {\textstyle \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}}\prod_{k=0}^{n-1} {\textstyle \Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}},
% \]
% so that
% \[
% 	\phi(\theta) = \lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=0}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}
% \]
is a fixed point whenever this limit exists.

Since $\varphi$ and $\Phi$ are characteristic functions, they are continuous and $\phi(0) = \Phi(0) = 1$, and thus both are non-zero in a neighbourhood of 0. In particular, $\psi(\theta) = \Log{\varphi}(\theta)$ and $\Psi(\theta) = \Log{\Phi(\theta)}$ exist in some neighbourhood of 0 as well\footnote{We use $\Log$ to denote the principal value of the complex logarithm, for which the imaginary part always lies in the interval $(-\pi,\pi]$.}.  From the above, we have
\[
	\Log{F^{(n)}(\varphi)(\theta)} = {\textstyle 2^{n} \psi\left(\frac{\theta}{2^{n}}\right)} 
		+ \sum_{k=0}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}.
\]
We briefly explore what possible limiting values are valid characteristic functions. 
We refer to the first term on the left hand side as the \emph{reproduction term} and the remaining sum as the \emph{noise term}, and only consider the case \revpoint{2}{8}
when both have well-posed limits independent of the other. \llabel{limit_remark} % We don't need to limit to this caee.
Biologically, \llabel{clarification2}
the ``noise term'' is the contribution to variation in the population
generated by the Mendelian sampling terms (\ie, by recombination and mutation),
while the ``reproduction term'' is the contribution from variation in the initial generation.

\paragraph{Noise Terms}
First, consider the noise terms.  A necessary condition for $\sum_{k=0}^{\infty} 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)$ to converge is that $2^{k} \Psi\left(\frac{\theta}{2^{k}}\right) \to 0$; since $\Psi(0) = 0$, this implies that if it exists, $\Psi'(0+) = 0$.  In particular, as we would expect, if $R$ has a mean, it must be 0.  If, on the other hand, $\Psi(\theta)$ is twice differentiable, so $R$ has mean 0 and finite variance, say $\Psi''(0) = \eta^{2}$, then Taylor's theorem tells us that for $\theta$ sufficiently small, 
\[
	\Psi(\theta) = \frac{\eta^{2}}{2} \theta^{2} + r(\theta)\theta^{2},
\]
where $r(\theta) \to 0$ as $\theta \to 0$.  Thus,
\begin{equation*}
	\frac{\textstyle 2^{k+1} \Psi\left(\frac{\theta}{2^{k+1}}\right)}{\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
	= \frac{1}{2} \frac{\eta^{2} \theta^{2} + {\textstyle  r\left(\frac{\theta}{2^{k+1}}\right)}\theta^{2}}{\eta^{2} \theta^{2} + {\textstyle  r\left(\frac{\theta}{2^{k}}\right)}\theta^{2}}
	\to \frac{1}{2},
\end{equation*}
as $k \to \infty$ and the sum converges by the ratio test.  In this case,
\begin{align*}
	 \sum_{k=0}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
	 &=  \sum_{k=0}^{n-1} 2^{k} {\textstyle \left(- \frac{\eta^{2}}{2}  \frac{\theta^{2}}{2^{2k}}
		+ r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{2k}}\right)}\\
	&=- \eta^{2} \theta^{2} (1 - 2^{-n+1})  + \sum_{k=0}^{n-1} {\textstyle r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{k}}}
\end{align*}
If we assume that $r(\theta) \equiv 0$, so that the noise is $N(0,\eta^{2})$, then the limiting distribution is $N(0,2\eta^{2})$. When  $r(\theta)$ is non-zero, we don't expect the higher order terms to vanish.

@@

\paragraph{Reproductive Terms:}
We now turn our attention to the limit
\[
		\alpha(\theta) = \lim_{n \to \infty} 2^{n}\psi\left(\frac{\theta}{2^{n}}\right)
\]
If it exists, then for all integers $k$, 
\[
	\alpha(2^{k} \theta) = \lim_{n \to \infty} 2^{n}\psi\left(\frac{2^{k} \theta}{2^{n}}\right) =  \lim_{n \to \infty} 2^{k} 2^{n-k} \psi\left(\frac{\theta}{2^{n-k}}\right) = 2^{k} \alpha(\theta).
\]
and, since $\alpha(0) = 0$,
\[
	\frac{\alpha(\theta)}{\theta} = \frac{\alpha(2^{k} \theta)}{2^{k}\theta} 
	\to \begin{cases} 
		\alpha'(0+) & \text{if $\theta > 0$, and}\\
		\alpha'(0-) & \text{if $\theta < 0$,}
	\end{cases}
\]
as $k \to - \infty$ if the limits exist (\ie there are unique left-and right-hand limits at zero; this is the case if \eg we require $\alpha(\theta)$, and thus $\phi(\theta)$, to be convex functions).  We henceforth assume this.

Since $\varphi(\theta)$ is a characteristic function, it has the Hermitian property: $\phi(-\theta) = \overline{\phi(\theta)}$, and thus $\psi(-\theta) = \overline{\psi(\theta)}$ and $\alpha(-\theta) = \overline{\alpha(\theta)}$.  In particular, if one (and by extension both) limits exist, we must have $\alpha'(0-) = \overline{\alpha'(0+)}$.  In this case, setting $-\gamma + i m = \alpha'(0+)$, we have
\[
	\alpha(\theta) = \begin{cases} 
		i m\theta - \gamma \theta & \text{if $\theta \geq 0$, and}\\
		i m\theta + \gamma \theta & \text{if $\theta < 0$}.
	\end{cases}
\]
If $\gamma > 0$, then we recognize $e^{\psi(\theta)} = e^{im \theta -\gamma|\theta|}$
as the characteristic function of a Cauchy distribution with median 0 and median absolute deviation $\gamma$
(\ie, the canonical $\alpha$-stable distribution with $\alpha=1$), \llabel{cauchy_note}
whereas if $\gamma = 0$, then $e^{\psi(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, \ie the characteristic function for a monomorphic population with all individuals having trait value equal to $m$.

% This might seem a bit roundabout, but the point is that we can prove that $e^{\psi(\theta)}$ is -- in this case -- a characteristic function, so we don't need to lose generality by assuming that it is one.  Our goal is to ask when $\lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=0}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}$ is a valid characteristic function, and to see what it can be.  
To exclude the case $\gamma < 0$, we first observe that if, as we assume, $\varphi(\theta)$ is the characteristic function of some random variable, say $X$, then $\varphi_{n}(\theta) = \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}} = e^{2^{n}\psi\left(\frac{\theta}{2^{n}}\right)}$ is the characteristic function of $2^{-n} \sum_{i = 1}^{2^{n}} X_{i}$, where $X_{1},\ldots,X_{2^{n}}$ are i.i.d.\ copies of $X$; now, L\'evy's continuity theorem (see \eg Theorem 2.3.4 in \cite{Durrett2005}) tells us that if (i) $\phi_{n}(\theta)$ has a point-wise limit, say $\phi_{\infty}(\theta)$, and (ii) $\phi_{\infty}(\theta)$ is continuous at 0, then $\phi_{\infty}(\theta)$ is the characteristic function of \emph{some} random variable; in particular, if $\alpha(\theta)$ exists and is bounded (\eg if the left and right-hand side derivatives exist, as above), then $\psi(\theta) = \alpha(\theta)\theta$ is continuous at 0, and hence so is $e^{\psi(\theta)}$.  Thus, in this case $e^{\psi(\theta)}$ is itself a characteristic function.  Moreover, being a characteristic function, $e^{\psi(\theta)}$ must satisfy $|e^{\psi(\theta)}| = e^{\Re \psi(\theta)} \leq 1$, whence $\gamma \geq 0$.

>>>>

The Gaussian distribution appears thanks to ``the'' Central Limit Theorem,
i.e., the universality of adding up a great many very small and independent things.
However, there is a wider class of Central Limit Theorems,
which describe the distributions of sums of a great many independent and exchangeable things:
the \emph{stable laws} \citep{gnedenko1968limit}.
Roughly speaking, these say that if we add together $n$ independent copies of a random variable $X$
for which $\P\{ X > t \}$ is proportional to $t^{-\alpha}$ for some $0 < \alpha < 2$,
then -- almost regardless of the details -- the result, scaled by $n^{1/\alpha}$,
is well-approximated by a universal form, an $\alpha$-stable distribution
(and the approximation becomes exact, as $n \to \infty$).
On the other hand, if $\alpha \ge 2$ then the distribution has finite variance, \revpoint{2}{1} 
and so the sum of $n$, divided by $\sqrt{n}$, converges to a Gaussian distribution.
In such ensembles, single large entries can be important:
concretely, the largest of the $n$ copies will be of order $n^{1/\alpha}$,
so that e.g., for $\alpha=1$ a positive proportion of the sum of $n$ entries will be contributed by a single one.
On the other hand, the largest of $n$ Gaussians is of order $\sqrt{\log n}$. \revpoint{1}{11}

@@

Since the Mendelian sampling term is a sum of terms of the form
$$
    (X_\ell - 1/2) \eta_\ell^{[1]} + (1/2 - X_\ell) \eta_\ell^{[2]},
$$
where $\eta_\ell^{[i]}$ is the effect of the allele in parent $i$ at locus $\ell$,
the appropriate central limit theorem for the sum is determined by
the tail behavior of $|(X - 1/2) \eta|$,
where $X$ is Bernoulli(1/2) and $\eta$ is the effect
of a uniformly chosen one of the two alleles from a uniformly chosen parent
at a uniformly chosen locus.
Since $|X-1/2| = 1/2$, we'd like to know
if $\P\{|\eta|>2t\}$ is proportional to $t^{-\alpha}$,
i.e., the value of
$$
    \lim_{t \to \infty} \frac{ - \log \P\{ |\eta| > 2 t \} }{ \log t } = \alpha ,
$$
if the limit exists.
Let $p_\ell$ be the minor allele frequency at the $\ell^\text{th}$ locus;
since the data we have assigns effect 0 to the major allele,
$\eta_\ell = 0$ with probability $1-p_\ell$.
Therefore, if we have estimated effects for $M$ SNPs
with minor allele frequency $p_\ell$ and effect size $e_\ell$ for the minor allele,
for $1 \le \ell \le M$,
then
$$
    \P\{ |\eta_{\ell}| > 2t \} \approx \frac{1}{M} \sum_{\ell : |e_\ell| > 2t} p_\ell .
$$
Note that the relevant effect size distribution here is weighted by frequency, \revpoint{1}{3}
which is necessary because we're looking at the deviation of a typical offspring from their parents,
and the chance a given allele is carried by a random parent is given by the frequency.
(This weighting will make a difference, as frequencies and effect sizes are often correlated in practice.)
So, to obtain a per-trait estimate of $\alpha$,
we plot $\log \frac{1}{M} \sum_{\ell : |e_\ell| > t} p_\ell$ against $\log t$
for a sequence of values of $t$,
and fit a linear model to the tail.
(To do this,
we take as the ``tail''
all values of $t$ for which the number of SNPs with absolute effect above $t$
is at least 20 and at most 10\% of all SNPs.) \revpoint{2}{3}
Examples are shown in Figure~\ref{fig:example_snps}.

@@

Concretely, consider a Moran process where an individual whose trait value is $Z$ dies at rate $\mu(Z)$
and is replaced by the offspring of two randomly chosen individuals,
i.e., a new individual with trait \llabel{offspring_eqn}
\begin{align*}
  \frac{Z^{1} +Z^{2}}{2} + R ,
\end{align*}
where $Z^{1}$ and $Z^{2}$ are independent draws from the empirical population distribution
and $R$ is an independent draw from the Mendelian sampling distribution
(independent of $Z^{1}$ and $Z^{2}$, with distribution to be specified). 
(Note that mutation might enter into $R$, but as the underlying genetic basis is not specified,
we do not attempt to disambiguate the various possible contributions.) \llabel{mutation_note}
Mutation-selection balance demands that that
the distribution of individuals removed by death
and the distribution of individuals added by birth are equal,
\ie that $(Z^{1} +Z^{2})/2 + R$ is equal in distribution
to a $\mu$-weighted choice of $Z$.
% In other words, for all bounded and continuous functions $f$,
% \begin{equation} \label{eqn:MSB_condition}
%     \E\left[ f\left(\frac{Z^{1} +Z^{2}}{2} + R \right) \right] = \frac{\E\left[ f(Z) \mu(Z) \right]}{\E[\mu(Z)]},
% \end{equation}
% where the expectation is with respect to the joint law of the population empirical distribution (the law of $Z$, $Z^{1}$ and $Z^{2}$), and $R$.
% 
% Since the set of functions $\mathcal{S} = \{e^{i \theta x} : \theta \in \R\}$ is separating (see \eg \S 3.4 in \cite{Ethier+Kurtz1986}), it suffices to show that \eqref{eqn:MSB_condition} holds for $f \in \mathcal{S}$, \ie
If the population is large enough, \llabel{deterministic_point}
we can treat the trait distribution as changing deterministically.
Taking the characteristic function of both these distributions,
deterministic mutation-selection balance is therefore equivalent to
\begin{align} \label{eqn:MSB_CF}
    \frac{\E\left[ e^{i \theta Z}\mu(Z) \right]}{\E[\mu(Z)]}
    =
    \E\left[ e^{i\theta\left(\frac{Z^{1} +Z^{2}}{2} + R \right)}\right] .
\end{align}
Now, we assumed that $Z^{1}$, $Z^{2}$, and $R$ are independent, so we can write \eqref{eqn:MSB_CF} as 
\begin{align*}
    \frac{\E\left[ e^{i\theta Z} \mu(Z) \right]}{\E[\mu(Z)]} &= \E\left[ e^{i\theta\left(\frac{Z^{1} +Z^{2}}{2} + R \right)}\right]\\
	% &= \E\left[ e^{i\frac{\theta}{2} Z^{1}}e^{i\frac{\theta}{2} Z^{2}}e^{\i\theta R}\right]\\
	&= \E\left[e^{i\frac{\theta}{2} Z^{1}}\right]\E\left[e^{i\frac{\theta}{2} Z^{2}}\right]\E\left[e^{i\theta R}\right]\\
	&= \phi\left(\frac{\theta}{2}\right)^{2} \Phi(\theta),
\end{align*}
where
$ \phi(\theta) = \mathbb{E}\left[e^{i\theta Z^{1}}\right] = \mathbb{E}\left[e^{i\theta Z^{2}}\right] $
and
$ \Phi(\theta) = \mathbb{E}\left[e^{i\theta R}\right] $
are the characteristic functions of $Z^{j}$ and $R$, respectively. 
In what follows, we will explore some non-Gaussian solutions $\phi(\theta)$ to equation \eqref{eqn:MSB_CF}.

@@

First, consider the case without selection, so $\mu$ is a constant
which we may take equal to 1. 
(To remind the reader, our goal is therefore to describe \llabel{clarification1}
what forms the population distribution might take in a neutral model
if we assume only that the Mendelian sampling terms are independent of parental values,
and nothing else about the distribution of these terms or the initial population distribution.) 
Then, our condition on the characteristic function simplifies to
\begin{equation}\label{eq:CF_neutral}
	{\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta) %e^{-\frac{\eta^{2}\theta^{2}}{2}} 
	= \phi(\theta) ,
\end{equation}
\ie $\phi(\theta)$ is a fixed point of $F$, where
 \[
	F(\phi)(\theta) = {\textstyle \phi\left(\frac{\theta}{2}\right)^{2}} \Phi(\theta),
\]
whereas any fixed point must be the limit of an arbitrary function, $\varphi(\theta)$, under repeated applications of $F$,
\ie \revpoint{2}{7}
\begin{equation}\label{eq:fpeq}
\begin{aligned}
	\phi(\theta) 
    &= \lim_{n \to \infty} F^{(n)}(\varphi)(\theta) \\
    &= \lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=0}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}} .
\end{aligned}
\end{equation}
% Iterating $F$ $n$ times, we get
% \[
% 	F^{(n)}(\varphi)(\theta) =  {\textstyle \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}}\prod_{k=0}^{n-1} {\textstyle \Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}},
% \]
% so that
% \[
% 	\phi(\theta) = \lim_{n \to \infty} \varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}\prod_{k=0}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}
% \]
is a fixed point whenever this limit exists, whereas any fixed point $\phi$ of $F$ satisfies  $\phi(\theta) = \lim_{n \to \infty} F^{(n)}(\phi)(\theta)$.  As we are interested in identifying those fixed points that are characteristic functions, we henceforth assume, without loss of generality, that the $\varphi$ is itself a characteristic function (if $\phi$ exists then we can always take $\varphi = \phi$) and that the limit $\lim_{n \to \infty} F^{(n)}(\varphi)(\theta)$ exists.

Since $\varphi$ and $\Phi$ are characteristic functions, say for real-valued random variables $Z$ and $R$, respectively, then $\varphi\left(\frac{\theta}{2^{n}}\right)^{2^{n}}$, $\Phi_{n}(\theta) = \prod_{k=0}^{n-1}\Phi\left(\frac{\theta}{2^{k}}\right)^{2^{k}}$, and $F^{(n)}(\varphi)(\theta)$ are all characteristic functions, of 
\[
	 2^{-n} \sum_{i = 1}^{2^{n}} Z_{i}, \quad \sum_{k = 0}^{n-1} 2^{-k} \sum_{j = 1}^{2^{k}} R_{j}, 
	 \quad \text{and} \quad 
	 2^{-n} \sum_{i = 1}^{2^{n}} Z_{i} + \sum_{k = 0}^{n-1} 2^{-k} \sum_{j = 1}^{2^{k}} R_{k,j},
\]
respectively, where the $Z_{i}$ and $R_{k,j}$ are i.i.d. copies of $Z$ and $R$.   L\'evy's continuity theorem (see \eg Theorem 2.3.4 in \cite{Durrett2005}) tells us that if $\varphi_{n}(\theta)$ has a point-wise limit, say $\varphi_{\infty}(\theta)$, then $\varphi_{\infty}(\theta)$ is the characteristic function of some random variable if and only if $\varphi_{\infty}(\theta)$ is continuous at 0 (in which case it is continuous for all $\theta \in \mathbb{R}$), and similarly for $\Phi_{n}(\theta)$ and $F^{(n)}(\varphi)(\theta)$.  Note that, since we are assuming that $F^{(n)}(\varphi)(\theta)$ has a point-wise limit, then either both $\varphi_{n}$ and $\Phi_{n}$ have point-wise limits, or neither does.

Now, as characteristic functions, $\varphi$ and $\Phi$ are continuous and $\varphi(0) = \Phi(0) = 1$, and thus both are non-zero in some neighbourhood of 0. In particular, $\psi(\theta) = \Log{\varphi}(\theta)$ and $\Psi(\theta) = \Log{\Phi(\theta)}$ exist in some neighbourhood of 0 as well\footnote{We use $\Log$ to denote the principal value of the complex logarithm, for which the imaginary part always lies in the interval $(-\pi,\pi]$.}.  From the above, we have
\[
	\Log{F^{(n)}(\varphi)(\theta)} = {\textstyle 2^{n} \psi\left(\frac{\theta}{2^{n}}\right)} 
		+ \sum_{k=0}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}.
\]
We briefly explore what possible limiting values are valid characteristic functions. 
We refer to the first term on the left hand side, $\psi_{n}(\theta) = 2^{n} \psi\left(\frac{\theta}{2^{n}}\right)$,  as the \emph{reproduction term} and the remaining sum, $\Psi_{n}(\theta) = \sum_{k=0}^{n-1} 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)$, as the \emph{noise term}, and only consider the case \revpoint{2}{8}
when both have well-posed limits independent of the other. \llabel{limit_remark} % We don't need to limit to this caee.
Biologically, \llabel{clarification2}
the ``noise term'' is the contribution to variation in the population
generated by the Mendelian sampling terms (\ie, by recombination and mutation),
while the ``reproduction term'' is the contribution from variation in the initial generation.

\paragraph{Noise Terms}
First, consider the noise terms.  A necessary condition for $\Psi_{n}(\theta)$ to converge is that $2^{k} \Psi\left(\frac{\theta}{2^{k}}\right) \to 0$; since $\Psi(0) = 0$, this implies that if it exists, $\Psi'(0+) = 0$.  In particular, as we would expect, if $R$ has a mean, it must be 0.  If, on the other hand, $\Psi(\theta)$ is twice differentiable, so $R$ has mean 0 and finite variance, say $\Psi''(0) = \eta^{2}$, then Taylor's theorem tells us that for $\theta$ sufficiently small, 
\[
	\Psi(\theta) = \frac{\eta^{2}}{2} \theta^{2} + r(\theta)\theta^{2},
\]
where $r(\theta) \to 0$ as $\theta \to 0$.  Thus,
\begin{equation*}
	\frac{\textstyle 2^{k+1} \Psi\left(\frac{\theta}{2^{k+1}}\right)}{\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
	= \frac{1}{2} \frac{\eta^{2} \theta^{2} + {\textstyle  r\left(\frac{\theta}{2^{k+1}}\right)}\theta^{2}}{\eta^{2} \theta^{2} + {\textstyle  r\left(\frac{\theta}{2^{k}}\right)}\theta^{2}}
	\to \frac{1}{2},
\end{equation*}
as $k \to \infty$ and the sum converges by the ratio test.  In this case,
\begin{align*}
	 \sum_{k=0}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
	 &=  \sum_{k=0}^{n-1} 2^{k} {\textstyle \left(- \frac{\eta^{2}}{2}  \frac{\theta^{2}}{2^{2k}}
		+ r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{2k}}\right)}\\
	&=- \eta^{2} \theta^{2} (1 - 2^{-n+1})  + \sum_{k=0}^{n-1} {\textstyle r\left(\frac{\theta}{2^{k}}\right)\frac{\theta^{2}}{2^{k}}}
\end{align*}
If we assume that $r(\theta) \equiv 0$, so that the noise is $N(0,\eta^{2})$, then the limiting distribution is $N(0,2\eta^{2})$. When  $r(\theta)$ is non-zero, we don't expect the higher order terms to vanish.

@@

\paragraph{Reproductive Terms:}
We now turn our attention to the (point-wise) limit
\[
		\psi_{\infty}(\theta) = \lim_{n \to \infty} \psi_{n}(\theta).
\]
Since $\varphi(\theta)$ is a characteristic function, it has the Hermitian property: $\varphi(-\theta) = \overline{\varphi(\theta)}$, and thus $\psi_{n}(-\theta) = \overline{\psi_{n}(\theta)}$ for all $n$ and $\psi_{\infty}(-\theta) = \overline{\psi_{\infty}(\theta)}$ for all $\theta$ at which the limit exists.  We can thus, without loss, assume that $\theta > 0$.

Suppose $\psi_{\infty}(\theta)$ exists for some $\theta > 0$.  Then for all integers $k$, 
\[
	\psi_{\infty}(2^{k} \theta) = \lim_{n \to \infty} 2^{n}\psi\left(\frac{2^{k} \theta}{2^{n}}\right) =  \lim_{n \to \infty} 2^{k} 2^{n-k} \psi\left(\frac{\theta}{2^{n-k}}\right) 
	= 2^{k} \psi_{\infty}(\theta).
\]
In particular, since $\psi_{\infty}(0) = 0$, 
\[
	\frac{\psi_{\infty}(\theta)}{\theta} = \frac{\psi_{\infty}(2^{k} \theta)}{2^{k}\theta} = \frac{\psi_{\infty}(2^{k} \theta)-\psi_{\infty}(0)}{2^{k}\theta - 0}, 
\]
so if $\psi_{\infty}$ has a right-hand side derivative at zero, then
\[
	\frac{\psi_{\infty}(\theta)}{\theta} = \lim_{k \to -\infty} \frac{\psi_{\infty}(2^{k} \theta)}{2^{k}\theta} = \psi_{\infty}'(0+).
\]

Suppose that this derivative exists, say $\psi_{\infty}'(0+) = -\gamma + i m$ for some $m, \gamma \in \mathbb{R}$.  Then, by the Hermitian property,  $\psi_{\infty}'(0-) = \overline{\psi_{\infty}'(0+)}$ must exist as well, so
\[
	\psi_{\infty}(\theta) = \begin{cases} 
		i m\theta - \gamma \theta & \text{if $\theta \geq 0$, and}\\
		i m\theta + \gamma \theta & \text{if $\theta < 0$}.
	\end{cases}
\]
Moreover,  $\psi_{\infty}(\theta) = i m \theta - \gamma |\theta|$ is continuous at 0, and hence so is $\varphi_{\infty}(\theta) = e^{\psi_{\infty}(\theta)}$.  As we observed above, this implies that $e^{\psi_{\infty}(\theta)}$ is itself a characteristic function.  Thus, $e^{\psi_{\infty}(\theta)}$ must satisfy $|e^{\psi_{\infty}(\theta)}| = e^{\Re \psi_{\infty}(\theta)} \leq 1$, whence $\Re \psi_{\infty}(\theta) \leq 0$ and $\gamma \geq 0$.  In particular, if $\gamma > 0$, then we recognize $e^{\psi_{\infty}(\theta)} = e^{im \theta -\gamma|\theta|}$ as the characteristic function of a Cauchy distribution with median m and median absolute deviation $\gamma$ (\ie, the canonical $\alpha$-stable distribution with $\alpha=1$), \llabel{cauchy_note} whereas if $\gamma = 0$, then $e^{\psi_{\infty}(\theta)} = e^{im \theta}$ is the Fourier transform of a Dirac mass at $m$, \ie the characteristic function for a monomorphic population with all individuals having trait value equal to $m$.
